Models of VOR learning in MHC knockout mice

### 2013-02-04 Notes/VORmodel.tex
simplified figures

We will be parsimonious in our modelling.
We will assume that the wild-type has a symmetry between potentiation and depression and we will treat gain-increase and decrease training in a symmetric way.
We will also assume that the rates of potentiation/depression in the mutant and wild-type are the same.
None of these assumptions are necessary or justified experimentally (although \cite{Fusi2007multistate} notes that imbalance between potentiation and depression can drastically impair memory).
They do reduce the number of adjustable parameters we have, simplifying searches for good models.
Presumably, relaxing these assumptions and enlarging the parameter space would allow for a better fit to data, but the number of parameters would start to look large compared to the number of features we're trying to match.

@@

Within the parameter range explored, we will see that synaptic complexity is essential.
The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary_res}), in contradiction with experiment.
It seems that the enhanced depressing transition more than compensates for saturation effects.

@@

All of these models show pre-training enhancing learning for the mutant, as expected.
It is harder to find regimes where pre-training impairs learning for the wild-type.
In the cascade model, this required a much longer duration for pe-training (see \autoref{fig:cascade_long}).
In the multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression.
For the pooled resource model, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.

@@

@@

We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition. We will only allow the synaptic weight to take two values, whereas it originally varied linearly across the states. } (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary}).
We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary}
 \end{myenuma}
 \end{center}
  \caption{Transition probabilities for different models.
  (\ref{fig:cascade}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
  (\ref{fig:multistate}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
\end{figure}

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
%  \begin{tabular}{|l|c|c|c|c|c|c|>{\refstepcounter{tablerow}}l}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50  &\label{tr:cascade_short} \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 &\label{tr:cascade_long} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50  &\label{tr:multistate_weak} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20  & \label{tr:multistate_strong} \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5   &\label{tr:binary} \\
    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

This compound synapse would seem to have $2^P$ internal states.
However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
%
\begin{equation}\label{eq:pooledweight}
  \w_i = \frac{2i}{P}-1.
\end{equation}
%

@@

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{cascade_short_learn.eps}}\label{fig:cascade_short_learn}
  \item\aligntop{\includegraphics[width=7cm]{cascade_short_learnS.eps}}\label{fig:cascade_short_learnS}
  \item\label{fig:cascade_short_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{cascade_short_eq_WT.eps}}\label{fig:cascade_short_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{cascade_short_eq_KO.eps}}\label{fig:cascade_short_eq_KO}
                  \end{myenumi}
  \item\label{fig:cascade_short_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_WT_nopre.eps}}\label{fig:cascade_short_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_WT_pre.eps}}\label{fig:cascade_short_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_KO_nopre.eps}}\label{fig:cascade_short_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_KO_pre.eps}}\label{fig:cascade_short_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$).
  Other parameters can be found in \autoref{tr:cascade_short} of \autoref{tab:params}.
  (\ref{fig:cascade_short_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:cascade_short_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:cascade_short_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_short_eq_WT}) wild-type and (\ref{fig:cascade_short_eq_KO}) \KO\ mutant.
  (\ref{fig:cascade_short_pr}) Evolution of probability distributions for (\ref{fig:cascade_short_pr_WT_nopre}\ref{fig:cascade_short_pr_WT_pre}) wild-type and  (\ref{fig:cascade_short_pr_KO_nopre}\ref{fig:cascade_short_pr_KO_pre}) \KO\ mutant without (\ref{fig:cascade_short_pr_WT_nopre},\ref{fig:cascade_short_pr_KO_nopre}) and with (\ref{fig:cascade_short_pr_WT_pre},\ref{fig:cascade_short_pr_KO_pre}) pre-training. } \label{fig:cascade_short}
\end{figure}

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{cascade_long_learn.eps}}\label{fig:cascade_long_learn}
  \item\aligntop{\includegraphics[width=7cm]{cascade_long_learnS.eps}}\label{fig:cascade_long_learnS}
  \item\label{fig:cascade_long_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{cascade_long_eq_WT.eps}}\label{fig:cascade_long_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{cascade_long_eq_KO.eps}}\label{fig:cascade_long_eq_KO}
                  \end{myenumi}
  \item\label{fig:cascade_long_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_WT_nopre.eps}}\label{fig:cascade_long_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_WT_pre.eps}}\label{fig:cascade_long_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_KO_nopre.eps}}\label{fig:cascade_long_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_KO_pre.eps}}\label{fig:cascade_long_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$).
  Other parameters can be found in \autoref{tr:cascade_long} of \autoref{tab:params}.
  (\ref{fig:cascade_long_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:cascade_long_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:cascade_long_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_long_eq_WT}) wild-type and (\ref{fig:cascade_long_eq_KO}) \KO\ mutant.
  (\ref{fig:cascade_long_pr}) Evolution of probability distributions for (\ref{fig:cascade_long_pr_WT_nopre}\ref{fig:cascade_long_pr_WT_pre}) wild-type and  (\ref{fig:cascade_long_pr_KO_nopre}\ref{fig:cascade_long_pr_KO_pre}) \KO\ mutant without (\ref{fig:cascade_long_pr_WT_nopre},\ref{fig:cascade_long_pr_KO_nopre}) and with (\ref{fig:cascade_long_pr_WT_pre},\ref{fig:cascade_long_pr_KO_pre}) pre-training. } \label{fig:cascade_long}
\end{figure}

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_weak_learn.eps}}\label{fig:multistate_weak_learn}
  \item\aligntop{\includegraphics[width=7cm]{multistate_weak_learnS.eps}}\label{fig:multistate_weak_learnS}
  \item\label{fig:multistate_weak_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{multistate_weak_eq_WT.eps}}\label{fig:multistate_weak_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{multistate_weak_eq_KO.eps}}\label{fig:multistate_weak_eq_KO}
                  \end{myenumi}
  \item\label{fig:multistate_weak_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_WT_nopre.eps}}\label{fig:multistate_weak_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_WT_pre.eps}}\label{fig:multistate_weak_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_KO_nopre.eps}}\label{fig:multistate_weak_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_KO_pre.eps}}\label{fig:multistate_weak_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption{Simulation results for multistate model with weak training ($\Delta f=-0.1$).
  Other parameters can be found in \autoref{tr:multistate_weak} of \autoref{tab:params}.
  (\ref{fig:multistate_weak_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:multistate_weak_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:multistate_weak_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:multistate_weak_eq_WT}) wild-type and (\ref{fig:multistate_weak_eq_KO}) \KO\ mutant.
  (\ref{fig:multistate_weak_pr}) Evolution of probability distributions for (\ref{fig:multistate_weak_pr_WT_nopre}\ref{fig:multistate_weak_pr_WT_pre}) wild-type and  (\ref{fig:multistate_weak_pr_KO_nopre}\ref{fig:multistate_weak_pr_KO_pre}) \KO\ mutant without (\ref{fig:multistate_weak_pr_WT_nopre},\ref{fig:multistate_weak_pr_KO_nopre}) and with (\ref{fig:multistate_weak_pr_WT_pre},\ref{fig:multistate_weak_pr_KO_pre}) pre-training. } \label{fig:multistate_weak}
\end{figure}

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_strong_learn.eps}}\label{fig:multistate_strong_learn}
  \item\aligntop{\includegraphics[width=7cm]{multistate_strong_learnS.eps}}\label{fig:multistate_strong_learnS}
  \item\label{fig:multistate_strong_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{multistate_strong_eq_WT.eps}}\label{fig:multistate_strong_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{multistate_strong_eq_KO.eps}}\label{fig:multistate_strong_eq_KO}
                  \end{myenumi}
  \item\label{fig:multistate_strong_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_WT_nopre.eps}}\label{fig:multistate_strong_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_WT_pre.eps}}\label{fig:multistate_strong_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_KO_nopre.eps}}\label{fig:multistate_strong_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_KO_pre.eps}}\label{fig:multistate_strong_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption{Simulation results for multistate model with strong training ($\Delta f=-0.4$).
  Other parameters can be found in \autoref{tr:multistate_strong} of \autoref{tab:params}.
  (\ref{fig:multistate_strong_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:multistate_strong_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:multistate_strong_eq}) Equilibrium distributions with/without gain-increase/decrease training for (\ref{fig:multistate_strong_eq_WT}) wild-type and (\ref{fig:multistate_strong_eq_KO}) \KO\ mutant.
  (\ref{fig:multistate_strong_pr}) Evolution of probability distributions for (\ref{fig:multistate_strong_pr_WT_nopre}\ref{fig:multistate_strong_pr_WT_pre}) wild-type and  (\ref{fig:multistate_strong_pr_KO_nopre}\ref{fig:multistate_strong_pr_KO_pre}) \KO\ mutant without (\ref{fig:multistate_strong_pr_WT_nopre},\ref{fig:multistate_strong_pr_KO_nopre}) and with (\ref{fig:multistate_strong_pr_WT_pre},\ref{fig:multistate_strong_pr_KO_pre}) pre-training. } \label{fig:multistate_strong}
\end{figure}

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{binary_learn.eps}}\label{fig:binary_learn}
  \item\aligntop{\includegraphics[width=7cm]{binary_learnS.eps}}\label{fig:binary_learnS}
  \item\label{fig:binary_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{binary_eq_WT.eps}}\label{fig:binary_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{binary_eq_KO.eps}}\label{fig:binary_eq_KO}
                  \end{myenumi}
  \item\label{fig:binary_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_WT_nopre.eps}}\label{fig:binary_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_WT_pre.eps}}\label{fig:binary_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_KO_nopre.eps}}\label{fig:binary_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_KO_pre.eps}}\label{fig:binary_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption{Simulation results for two-state model.
  Parameters can be found in \autoref{tr:binary} of \autoref{tab:params}.
  (\ref{fig:binary_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:binary_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:binary_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:binary_eq_WT}) wild-type and (\ref{fig:binary_eq_KO}) \KO\ mutant.
  (\ref{fig:binary_pr}) Evolution of probability distributions for (\ref{fig:binary_pr_WT_nopre}\ref{fig:binary_pr_WT_pre}) wild-type and  (\ref{fig:binary_pr_KO_nopre}\ref{fig:binary_pr_KO_pre}) \KO\ mutant without (\ref{fig:binary_pr_WT_nopre},\ref{fig:binary_pr_KO_nopre}) and with (\ref{fig:binary_pr_WT_pre},\ref{fig:binary_pr_KO_pre}) pre-training. } \label{fig:binary_res}
\end{figure}

The results of simulations of the two-state model can be seen in \autoref{fig:binary_res}.
However, this model can be solved exactly:
%
\begin{multline}\label{eq:binarysol}
  \eq = \frac{(f\dep q\dep, f\pot q\pot)}{\lambda},
  \qquad
  \pr(t) = \eq + (\pr(t)-\eq)\e^{-\lambda rt},\\
  \qquad \text{where} \quad
  \lambda = f\pot q\pot + f\dep q\dep.
\end{multline}
%
But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
In this case, the initial rate of change encapsulates the whole solution, as there is only a single exponential decay.

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:binWTnopre}
  \Phi = -\Delta f q,
\end{equation}
%
where, again, it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:binWTpre}
\begin{aligned}
  \Phi &= -2\Delta f q.
\end{aligned}
\end{equation}
%
So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
This can be seen in \autoref{fig:binary_res}\ref{fig:binary_learnS}.

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:binKOnopre}
  \Phi = -\frac{2\Delta f q}{1+\beta},
\end{equation}
%
which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
This is also seen in \autoref{fig:binary_res}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
The second effect is much smaller for the very short chain considered here.

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_learn.eps}}\label{fig:pooled_plenty_learn}
  \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_learnS.eps}}\label{fig:pooled_plenty_learnS}
  \item\label{fig:pooled_plenty_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_eq_WT.eps}}\label{fig:pooled_plenty_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_eq_KO.eps}}\label{fig:pooled_plenty_eq_KO}
                  \end{myenumi}
  \item\label{fig:pooled_plenty_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_WT_nopre.eps}}\label{fig:pooled_plenty_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_WT_pre.eps}}\label{fig:pooled_plenty_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_KO_nopre.eps}}\label{fig:pooled_plenty_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_KO_pre.eps}}\label{fig:pooled_plenty_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$).
  Parameters can be found in \autoref{tr:pooled_plenty} of \autoref{tab:params}.
  (\ref{fig:pooled_plenty_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:pooled_plenty_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:pooled_plenty_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:pooled_plenty_eq_WT}) wild-type and (\ref{fig:pooled_plenty_eq_KO}) \KO\ mutant.
  (\ref{fig:pooled_plenty_pr}) Evolution of probability distributions for (\ref{fig:pooled_plenty_pr_WT_nopre}\ref{fig:pooled_plenty_pr_WT_pre}) wild-type and  (\ref{fig:pooled_plenty_pr_KO_nopre}\ref{fig:pooled_plenty_pr_KO_pre}) \KO\ mutant without (\ref{fig:pooled_plenty_pr_WT_nopre},\ref{fig:pooled_plenty_pr_KO_nopre}) and with (\ref{fig:pooled_plenty_pr_WT_pre},\ref{fig:pooled_plenty_pr_KO_pre}) pre-training. } \label{fig:pooled_plenty}
\end{figure}

>>>>

We will be parsimonious in our modelling.
We will assume that the wild-type has a symmetry between potentiation and depression and we will treat gain-increase and decrease training in a symmetric way.
We will also assume that the relative rates of potentiation/depression in the mutant and wild-type are the same and that the overall rate of plasticity events is the same in all situations.
None of these assumptions are necessary or justified experimentally (although \cite{Fusi2007multistate} notes that imbalance between potentiation and depression can drastically impair memory).
They do reduce the number of adjustable parameters we have, simplifying searches for good models.
Presumably, relaxing these assumptions and enlarging the parameter space would allow for a better fit to data, but the number of parameters would start to look large compared to the number of features we're trying to match.

@@

Within the parameter range explored, we will see that synaptic complexity is essential.
The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary}), in contradiction with experiment.
It seems that the enhanced depressing transition more than compensates for saturation effects.

@@

All of these models show pre-training enhancing learning for the mutant, as expected.
It is harder to find regimes where pre-training impairs learning for the wild-type.
In the cascade model, this required a much longer duration for pe-training (see \autoref{fig:cascade_long}).
In the multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression, or by allowing the rate of plasticity events to vary.
For the pooled resource model, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.

@@

We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equlibrium weights mentioned above.

@@

We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition. We will only allow the synaptic weight to take two values, whereas it originally varied linearly across the states. } (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary_model}).
We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
 \end{myenuma}
 \end{center}
  \caption{Transition probabilities for different models.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
\end{figure}

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50  &\label{tr:cascade_short} \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 &\label{tr:cascade_long} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50  &\label{tr:multistate_weak} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20  & \label{tr:multistate_strong} \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5   &\label{tr:binary} \\
    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

This compound synapse would seem to have $2^P$ internal states.
However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
%
\begin{equation}\label{eq:pooledweight}
  \w_i = \frac{2i}{P}-1.
\end{equation}
%

@@

\newcommand{\resultsfig}[2]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
  \item\label{fig:#1_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
                  \end{myenumi}
  \item\label{fig:#1_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#2]{#2.
  Other parameters can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training. } \label{fig:#1}
\end{figure}}

@@

\resultsfig{cascade_short}{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$)}

\resultsfig{cascade_long}{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$)}

@@

\resultsfig{multistate_weak}{Simulation results for multistate model with weak training ($\Delta f=-0.1$)}

\resultsfig{multistate_strong}{Simulation results for multistate model with strong training ($\Delta f=-0.4$)}

@@

\resultsfig{binary}{Simulation results for two-state model}

The results of simulations of the two-state model can be seen in \autoref{fig:binary}.
However, this model can be solved exactly:
%
\begin{multline}\label{eq:binarysol}
  \eq = \frac{(f\dep q\dep, f\pot q\pot)}{\lambda},
  \qquad
  \pr(t) = \eq + (\pr(t)-\eq)\e^{-\lambda rt},\\
  \qquad \text{where} \quad
  \lambda = f\pot q\pot + f\dep q\dep.
\end{multline}
%
But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
In this case, the initial rate of change encapsulates the whole solution, as there is only a single exponential decay.

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:binWTnopre}
  \Phi = -\Delta f q,
\end{equation}
%
where, again, it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:binWTpre}
\begin{aligned}
  \Phi &= -2\Delta f q.
\end{aligned}
\end{equation}
%
So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS}.

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:binKOnopre}
  \Phi = -\frac{2\Delta f q}{1+\beta},
\end{equation}
%
which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
The second effect is much smaller for the very short chain considered here.

@@

\resultsfig{pooled_plenty}{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}

\resultsfig{pooled_scarce}{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}
